---
title: Additional Taxi-out Time
categories:
  - metadata
  - glossary
---

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('static', 'css', 'table.css'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Contacts

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'contacts.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Metadata update

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'metadata-update.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Statistical presentation

### Data description
The **additional taxi-out time** is a proxy for the average departure runway queuing time
on the outbound traffic flow, during congestion periods at airports.

It is the difference between the actual taxi-out time of a flight and a statistically
determined reference taxi-out time based on taxi-out times in periods of
low traffic demand. 
 
Uncertainty of take-off clearance time and aircraft arriving time at runway holding stop bars
makes traffic supply to runways a stochastic phenomenon.
In order to ensure continuous traffic demand at runways and maximise runway usage,
a minimum level of queuing is required.
However, additional time is detrimental to taxi-out efficiency, fuel consumption and environment.
Therefore, there exists a trade-off between taxi-out efficiency and runway throughput.

When monitoring taxi-out performance at airports, it is to be stressed that the goal is
not to reduce taxi time to an unimpeded time - as this could negatively impact
on runway throughput - but rather to reduce additional taxi-out time and associated
fuel burn to the strict minimum. 

As an output of the ATMAP Group, additional taxi-out time has been in use as a commonly
agreed proxy for airport inefficiency in the taxi-out phase since 2008 and is compliant
with the [PI][pi] definition in the EU legislation.


### Classification system
Additional taxi-out time is classified per Member State, with a breakdown for each airport
subject to performance monitoring within the SES performance scheme.

### Sector coverage

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'sector-coverage.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```

### Statistical concepts and definitions
The **Taxi-out time** is defined as the time spent by a flight between its
actual off-block time (AOBT) and actual take-off time (ATOT).

The [Unimpeded taxi-out time][utaxi] is an average taxi-out time when there is no congestion.
There is one unimpeded time by departure airport, departure runway and departure stand.

The **Additional taxi-out time** is the difference between the actual taxi-out time and
the unimpeded taxi-out time.  


### Statistical unit
The statistical unit is the airport. Airport level data is also aggregated to States.

### Statistical population
The statistical population is the set of airports subject to performance monitoring within the
SES performance scheme.

### Reference area
The reference area is the [Single European Sky Area][ses-area]. 

### Time coverage
2011 is the first year for which data is presented.

### Base period
Not applicable.

## Unit of Measure
The additional taxi-out time is measured in minutes per [IFR][ifr] departure [min/dep].

## Reference period

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'reference-period.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Institutional Mandate

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'institutional-mandate.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


## Confidentiality

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'confidentiality.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Release policy

### Release calendar
Taxi-out additional time is released monthly with yearly aggregates.

### Release calendar access
Not applicable.

### User access
Information is disseminated to the general public via the [SES Data Portal][ses-portal]. 


## Frequency of dissemination
Data is published monthly with the annual performance aggregate being available in January
of the following year. 

## Dissemination format
Information is disseminated to the general public via the [SES Data Portal][ses-portal]. 

## Accessibility of documentation

### Documentation on methodology
As per the pertaining regulations (see [Institutional Mandate](#institutional-mandate)).<br>
For technical questions not addressed in [Statistical processing](#statistical-processing)
please contact: [PRU-Support@eurocontrol.int](mailto:PRU-Support@eurocontrol.int).<br>
Additional definitions of the terms used in the frame of the this [KPI][kpi]
are available in the [Metadata page](..).

### Quality documentation
There is no specific documentation on procedures applied for quality management
and quality assessment.

## Quality Management
Although data providers are responsible for data quality,
the [EUROCONTROL Performance Review Unit][pru] performs data validation and quality checks.

### Quality assurance
Data validation is performed by the Central Office for Delay Analysis (CODA), on behalf of PRU, on each data delivery by airports,
and data validation report are returned to the data providers.

### Quality assessment
Raw data is cross-checked with various sources ([Network manager][nm-cfmu], ANSP's,
airport operators, airport coordinators and air carriers).
A quality threshold is established, and data that does not pass the quality threshold is rejected.

If a field is found to be blank, it is tried to fill the missing value from an alternative
data source (i.e. a missing aircraft type in the airport data flow can be filled
with information from the [Network manager][nm-cfmu].) 


### Completeness
The data is collected for all airports subject to performance monitoring within
the SES performance scheme.
Data completeness is determined each month as per the above-mentioned quality checks.
Any missing data is reported to the provider. 


## Relevance
The information is published for performance monitoring purposes in accordance with
the relevant EU legislation. 


## Accuracy and reliability

### Overall accuracy
The accuracy of the measure is influenced by the availability of the stand/runway configuration
and the type of AOBT recording at the airports (manual vs. automated).


### Sampling error
There is no sampling and therefore no sampling error. 


## Timeliness and punctuality
### Timeliness
The information is published each month - in general around 30 days after the end
of the month in question.

###  Punctuality
The internal databases are updated daily. The [statistical processing](#statistical-processing)
is performed once per month.

## Comparability
### Comparability — geographical

The data is collected centrally by the [EUROCONTROL Performance Review Unit (PRU)][pru]
with delegation to the Central Office for Delay Analysis (CODA) and computed consistently for all airports
subject to performance monitoring within the SES performance scheme.

The interpretation of the measure and comparisons across airports require due consideration
of prevailing local circumstances (airport infrastructure, etc.) 


### Comparability over time
Comparisons over time are valid.

## Coherence
### Coherence — cross domain
Checks have been carried out with a number of airports and there is generally a good level
of coherence between the indicator results and the results from performance measurement systems
of airport operators. 

### Coherence — internal
Data is fully coherent from an internal perspective.

## Cost and burden
Not available. 


## Data revision
Subject to changes (i.e. infrastructure), there might be a need to change
unimpeded times accordingly.

## Statistical processing

### Source data
In accordance with EU legislation, the data is collected centrally for all the airports
subject to performance monitoring within the SES performance scheme.

The input variables used for the calculation are detailed in the 
[Statistical concepts and definitions](#statistical-concepts-and-definitions) section. 


### Frequency of data collection
The data is collected on a daily basis and transmitted for statistical processing
on a monthly basis. 


### Data collection
The data is collected by the [EUROCONTROL Performance Review Unit (PRU)][pru] and the
Central Office for Delay Analysis (CODA).


### Data validation
The data is validation as described in the [Quality Management](#quality-management) section.


### Data compilation 

#### Calculation of the indicator

Let 

* $c$, a combination of departure runway and group of stands, as described in the
  [methodology for unimpeded taxi-out time][utaximet],
* $f(c)$ a flight characterised by a combination $c$,
* $\mathrm{AcTXOT}(f(c))$ the actual taxi-out time for a flight $f(c)$, i.e. the elapsed time
  between the off-block time (AOBT) of the flight $f(c)$ and its actual take-off time (ATOT),
* $\mathrm{UTXOT}(c)$ the [unimpeded taxi-out time][utaxi] for a combination $c$.

The additional taxi-out time $\mathrm{AdTXOT}(f(c))$ is calculated for each flight $f(c)$
as the difference between the actual taxi-out time $\mathrm{AcTXOT}(f(c))$ of the flight
and the [unimpeded taxi-out time][utaxi] $\mathrm{UTXOT}(c)$ :

$$
\mathrm{AdTXOT}(f(c)) = \mathrm{AcTXOT}(f(c)) - \mathrm{UTXOT}(c)
$$

The additional taxi-out time $\mathrm{AdTXOT}(c)$ for a given combination $c$ is the average
of the additional taxi-out time $\mathrm{AdTXOT}(f(c))$ of all the flight $f(c)$
characterised by that combination $c$.

The additional taxi-out time $\mathrm{AdTXOT}$ for a given airport is the weighted average
of the additional taxi-out time $\mathrm{AdTXOT}(c)$, for all the combinations $c$
at that airport with their probability of occurrence.


### Adjustment
Corrections may be made when issues are discovered as described in the
[Quality Assessment](#quality-assessment) section of this document. 


## Comment
No comments.

### Disclaimer {.disclaimer}

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'disclaimer.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


[ansp]: https://ansperformance.eu/acronym/ansp/ "ANSP"
[atfm-delay]: https://ansperformance.eu/definition/atfm-delay/ "ATFM Delay"
[air-stats]: https://ansperformance.eu/definition/air-transport-statistics/ "Air Transport Statistics"
[ctot]: https://ansperformance.eu/acronym/ctot/ "CTOT"
[etot]: https://ansperformance.eu/acronym/etot/ "ETOT"
[fab]: https://ansperformance.eu/acronym/fab/ "FAB"
[fmp]: https://ansperformance.eu/acronym/fmp/ "FMP"
[kpi]: https://ansperformance.eu/acronym/kpi/ "KPI"
[ifr]: https://ansperformance.eu/acronym/ifr/ "IFR"
[law2010/691]: https://eur-lex.europa.eu/legal-content/EN/TXT/?uri=CELEX:32010R0691 "Commission Regulation (EU) No 691/2010"
[law2013/390]: https://eur-lex.europa.eu/legal-content/EN/TXT/?uri=CELEX:32013R0390 "Commission Implementing Regulation (EU) No 390/2013"
[law2019/317]: https://eur-lex.europa.eu/legal-content/EN/TXT/?uri=CELEX:32019R0317 "Commission Implementing Regulation (EU) 2019/317"
[nm-cfmu]: https://ansperformance.eu/acronym/nm-cfmu/ "NM"
[pi]: https://ansperformance.eu/acronym/pi/ "PI"
[pru]: https://ansperformance.eu/about/us/ "PRU"
[refloc]: https://ansperformance.eu/definition/reference-location/ "Reference Location"
[regu]: https://ansperformance.eu/definition/regulation/ "Regulation"
[ses-area]: https://ansperformance.eu/definition/ses-area/ "SES Area"
[ses-portal]: https://www.eurocontrol.int/prudata/dashboard/ "SES Data Portal"
[uasma]: https://ansperformance.eu/definition/unimpeded-asma-time/ "Unimpeded ASMA Time"
[uasmamet]: https://ansperformance.eu/methodology/unimpeded-asma-time/ "Unimpeded ASMA Time - Methodology"
[utaxi]: https://ansperformance.eu/definition/unimpeded-taxi-out-time/ "Unimpeded Taxi-out Time"
[utaximet]: https://ansperformance.eu/methodology/unimpeded-taxi-out-time/ "Unimpeded Taxi-out Time - Methodology"


