---
title: Airport ATFM Delay
categories:
  - metadata
  - glossary
---

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('static', 'css', 'table.css'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Contacts

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'contacts.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


## Metadata update

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'metadata-update.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


## Statistical presentation

### Data description

::: {.note}
Please note that software release 20.0 of the Network Manager on 04 April 2016 introduced a
change in the computation of ATFM delay for operational purposes.
ATFM delays on this web page are based on this new computation from 04 April 2016 onwards.
See the Source data section below for more information on the changes in NM release 20.0.
:::

The measure provides an indication of [ATFM delays][atfm-delay] on the ground due to constraints
at airports. 


In Europe, when traffic demand is anticipated to exceed the available capacity in en route centres
or at airports, Air Traffic Control (ATC) units may request the local
Flow Management Position ([FMP][fmp]) to instigate an Air Traffic Flow Management (ATFM) measure,
or regulation.
Aircraft expected to arrive during a period of congestion are given Air Traffic Flow Management
delay at their departure airport, under the authority of the [Network Manager][nm-cfmu],
in order to regulate the flow of traffic into the constrained downstream en route sector or airport,
thus ensuring safety.
The resulting ATFM delays are calculated as the difference between the estimated take-off time
calculated from the filed flight plan including updates and the calculated take-off time
allocated by the central unit of ATFM.
The reason for the regulation is indicated by the responsible Flow Management Position (FMP).
The delay is attributed to the most constraining ATC unit.   

As an output of the ATMAP Group, Airport ATFM delay has been in use as a commonly agreed proxy
for airport capacity shortfalls since 2009.
The calculation of Airport ATFM delay is based on a well established and commonly accepted
algorithm and is compliant with the [KPI][kpi] definition in the EU legislation. 


### Classification system
Airport ATFM delay is classified per Member State, with a breakdown by airport.


### Sector coverage

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'sector-coverage.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


### Statistical concepts and definitions

ATFM delay
: *ATFM delay* is defined as the duration between the last Estimated Take-Off Time ([ETOT][etot])
  and the Calculated Take-Off Time ([CTOT][ctot]) allocated by the [Network Manager][nm-cfmu].
  ATFM delay comprises both Airport ATFM delay] and En route ATFM delay.
  This indicator is solely concerned with En route ATFM delay.
:  The delay is truncated to minutes.

Flights 
: All the [IFR][ifr] flights which have filed a flight plan and have landed at the relevant
  airport are taken into account.

#### Attribution of Delay
[ATFM delay][atfm-delay] to which a flight is subjected is defined to have as cause
the most penalizing [ATFM regulation][regu] which affects it.
Depending of the type and the entity responsible for the [reference location][refloc] protected by
the [ATFM regulation][regu] the [ATFM delay][atfm-delay] is classified as  either an airport or
an en route [ATFM delay][atfm-delay] and attributed to the  ANSP (en route) or airport.
Airport ATFM delays generated by arrival or global regulations are attributed to the
arrival airport as stated in the flight plan.
If the regulation is protecting a group of airports, the delay is attributed to the
destination airport of the affected flight. 

Each ATFM regulation also comprises a designator referring to the cause of delay
which can be used in further analysis.



### Statistical unit
The statistical unit is the airport. Airport level data is also aggregated to States.

### Statistical population
The statistical population is the set of airports subject to performance monitoring within
the Single European Sky Performance Scheme. 

### Reference area

The reference area is the [Single European Sky Area][ses-area]. 


### Time coverage
2009 is the first year for which data is presented.

### Base period
Not applicable.


## Unit of Measure
Airport [ATFM delay][atfm-delay] is measured in minutes per [IFR][ifr] arrival. 


## Reference period

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'reference-period.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```



## Institutional Mandate

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'institutional-mandate.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


## Confidentiality

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'confidentiality.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


## Release policy

### Release calendar
Airport [ATFM delay][atfm-delay] data is released monthly with yearly aggregates. 

### Release calendar access
Not applicable.

### User access
Information is disseminated to the general public via the [SES Data Portal][ses-portal]. 



## Frequency of dissemination
Airport ATFM data is published monthly with the annual performance aggregate
being available in January of the following year.


## Dissemination format
Information is disseminated to the general public via the [SES Data Portal][ses-portal]. 

## Accessibility of documentation

### Documentation on methodology
As per the pertaining regulations (see [Institutional Mandate](#institutional-mandate)).<br>
For technical questions not addressed in [Statistical processing](#statistical-processing)
please contact: [PRU-Support@eurocontrol.int](mailto:PRU-Support@eurocontrol.int).<br>
Additional definitions of the terms used in the frame of the this [KPI][kpi]
are available in the [Metadata page](..).

### Quality documentation
Documentation on procedures applied for quality management and quality assessment
is under development.

## Quality Management

### Quality assurance
The data is provided by the [Network Manager][nm-cfmu] and as such it is their
responsibility to perform any quality assurance activities.

### Quality assessment

Data is validated by random quality checks, e.g. missing data for given a day,
adherent data, etc.  Any issues are reported to the [Network manager][nm-cfmu] who works with the
[EUROCONTROL Performance Review Unit][pru] to resolve them.


## Relevance

The information is published for performance monitoring purposes in accordance with
the relevant EU legislation. 

### Completeness
The data is collected for the entire [Single European Sky Area][ses-area].

## Accuracy and reliability

### Overall accuracy
The Airport [ATFM delay][atfm-delay] is the delay calculated as defined in SES legislation. 
This calculation is based on a well established methodology, which has been in place since 2009. 


## Timeliness and punctuality
### Timeliness

The full set of Airport [ATFM delay][atfm-delay] per IFR arrival at airports
included in the performance scheme is published each month - in general between
14 and 16 days after the end of the month in question.


###  Punctuality
The internal databases are updated daily. The [statistical processing](#statistical-processing)
is performed once per month.

## Comparability
### Comparability — geographical
The data is collected centrally by the [Network Manager][nm-cfmu] and
computed consistently throughout the entire [Single European Sky Area][ses-area]
which allows comparisons across different airports.

### Comparability over time
Comparisons over time are valid.

## Coherence
### Coherence — cross domain
[ATFM delay][atfm-delay] is calculated relative to the flight plan. 

Although not fully cohered, the delay reported by aircraft operators to the Central Office for Delay Analysis (CODA)
under the corresponding IATA delay categories shows a high correlation with the ATFM delay
calculated by the Network Manager. 
It should be noted that the delay reported by aircraft operators relates to airline schedules
rather than flight plan.

### Coherence — internal
Data is fully coherent from an internal perspective.

## Cost and burden

Not available. 


## Data revision

The data revision policy has not yet been defined. 

## Statistical processing

### Source data

The data for the calculation of this indicator is provided by the [Network Manager][nm-cfmu].

> Please note that the delays caused by READY to Depart (REA) messages and ATFM slot extensions
> were excluded from the ATFM delay computation with NM release 20 (4 April 2016) which reduces
> the overall amount of computed ATFM delay.
> More information on the adjusted ATFM delay computation as of NM release 20.0
> from 4 April can be found here: [NM release 20.0 - 4 April 2016 - FB750-Delay Calculation improvements][nm20].


### Frequency of data collection
Data is collected for each individual flight in real time.


### Data collection

Data is collected and compiled by the [Network Manager][nm-cfmu].


### Data validation

Data is validated and cross-checked by the [Network Manager][nm-cfmu].


### Data compilation 

Data is aggregated for different delay groups, airports, and Member States. 

### Adjustment
Corrections may be made when issues are discovered as described in the
[Quality Assessment section of this document](#quality-assessment).

There are no systematic adjustments applied to this data.

## Comment
No comments.

### Disclaimer {.disclaimer}

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'disclaimer.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


[atfm-delay]: https://ansperformance.eu/definition/atfm-delay/ "ATFM Delay"
[ses-area]: https://ansperformance.eu/definition/ses-area/ "SES Area"
[kpi]: https://ansperformance.eu/acronym/kpi/ "KPI"
[ifr]: https://ansperformance.eu/acronym/ifr/ "IFR"
[fir]: https://ansperformance.eu/acronym/fir/ "FIR"
[fmp]: https://ansperformance.eu/acronym/fmp/ "FMP"
[nm-cfmu]: https://ansperformance.eu/acronym/nm-cfmu/ "NM"
[ansp]: https://ansperformance.eu/acronym/ansp/ "ANSP"
[fab]: https://ansperformance.eu/acronym/fab/ "FAB"
[air-stats]: https://ansperformance.eu/definition/air-transport-statistics/ "Air Transport Statistics"
[etot]: https://ansperformance.eu/acronym/etot/ "ETOT"
[ctot]: https://ansperformance.eu/acronym/ctot/ "CTOT"
[regu]: https://ansperformance.eu/definition/regulation/ "Regulation"
[refloc]: https://ansperformance.eu/definition/reference-location/ "Reference Location"
[law]: https://ansperformance.eu/bibliography/legislation/ "Legislation"
[pru]: https://ansperformance.eu/about/us/ "PRU"
[nm20]: https://ansperformance.eu/library/NM_20_FB750_Delay_Calculation_improvements.pdf "REA Message"
[ses-portal]: https://www.eurocontrol.int/prudata/dashboard/ "SES Data Portal"

