---
title: Share of en-route ATFM delay occurring when daily throughput was above expected daily traffic
categories:
  - metadata
  - glossary
---

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('static', 'css', 'table.css'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Contacts

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'contacts.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


## Metadata update

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'metadata-update.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```

### Indicator definition

The percentage of the total en route ATFM delay minutes that occurred on days when the daily 
throughput was above the expected daily traffic, calculated as follows:

* en route [ATFM delays][atfm-delay] calculated in accordance with point 3.1 of Section 1 of [Commission Implementing Regulation (EU) 2024/3128][no3128_2024] ;
* expected daily traffic refers to the daily traffic profile (guideline traffic) calculated by the Network Manager in January of each year;
* airspace throughput and traffic demand are measured as daily IFR flight entries in the airspace;
* the airspace is the area of responsibility defined for the calculation of en route ATFM delays.

## Unit of Measure

% of en route [ATFM delay][atfm-delay] on days when daily throughput was above the expected daily traffic.

### Statistical concepts and definitions

ATFM delay
: *ATFM delay* is defined as the duration between the last Estimated Take-Off Time ([ETOT][etot])
  and the Calculated Take-Off Time ([CTOT][ctot]) allocated by the [Network Manager][nm-cfmu].
  ATFM delay comprises both Airport ATFM delay] and En route ATFM delay.
  This indicator is solely concerned with En route ATFM delay.
:  The delay is truncated to minutes.

Flights 
: All the [IFR][ifr] flights which have filed a flight plan and have entered the area of
  responsibility of the relevant [ACC][acc](s) are taken into account.
: Each flight is only counted once, regardless of its trajectory.

#### Attribution of Delay

[ATFM delay][atfm-delay] to which a flight is subjected is defined to have as cause
the most penalizing [ATFM regulation][regu] which affects it.
Depending of the type and the entity responsible for the [reference location][refloc] protected by
the [ATFM regulation][regu] the [ATFM delay][atfm-delay] is classified as  either an airport or
an en route [ATFM delay][atfm-delay] and attributed to airport or en route facilities.

Each ATFM regulation also comprises a designator referring to the cause of delay
which can be used in further analysis.

### Statistical unit

The statistical unit is the [ACC][acc] to which the ATFM delays are attributed.

### Statistical population

The statistical population comprises the [ACC][acc] in the [Single European Sky Area][ses-area].

### Reference area

The reference area is the Single European Sky Area. 

### Time coverage

2025 is the first year for which data is presented.

## Reference period
```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'reference-period.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Institutional Mandate

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'institutional-mandate.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Release policy

### Release calendar

The indicator is calculated by the Network Manager in January of each year.
For information sharing purposes, the data is released 12 times per year, once per month.

### User access

Information is disseminated to the general public via the [SES Data Portal][ses-portal]. 

## Frequency of dissemination

The data is published monthly with the annual performance aggregate being available
in January of the following year.

## Dissemination format

Information is disseminated to the general public via the [SES Data Portal][ses-portal]. 

## Accessibility of documentation

### Documentation on methodology

As per the pertaining regulations (see [Institutional Mandate](#institutional-mandate)).<br>
For technical questions not addressed in [Statistical processing](#statistical-processing)
please contact: [PRU-Support@eurocontrol.int](mailto:PRU-Support@eurocontrol.int).<br>
Additional definitions of the terms used in the frame of the this [KPI][kpi]
are available in the [Metadata page](..).

### Quality documentation

There is no specific documentation on procedures applied for quality management
and quality assessment.

## Quality Management

### Quality assurance

The data is provided by the [Network Manager][nm-cfmu] and as such it is their
responsibility to perform any quality assurance activities.

### Quality assessment

Data is validated by random quality checks, e.g. missing data for given a day,
adherent data, etc.  Any issues are reported to the [Network manager][nm-cfmu] who works with the
[EUROCONTROL Performance Review Unit][pru] to resolve them.

## Relevance

The information is published for performance monitoring purposes in accordance with
the relevant EU legislation. 

### Completeness

The data is collected for the entire [Single European Sky Area][ses-area].

## Accuracy and reliability

### Overall accuracy

The en route [ATFM delay][atfm-delay] is the delay calculated by the [Network Manager][nm-cfmu].
This calculation is based on a well established methodology, which has been in place since 1999.

It is assumed that the data provided is accurate, given the caveats mentioned in sections
[Quality assurance](#quality-assurance) and
[Quality assessment](#quality-assessment).

### Sampling error

There is no sampling and therefore no sampling error. 

## Timeliness and punctuality

### Timeliness

The data is published each month
- in general between 14 and 16 days after the end of the month in question.

###  Punctuality

The internal databases are updated daily. The [statistical processing](#statistical-processing)
is performed once per month.

## Comparability
### Comparability — geographical

The data is collected centrally by the [Network Manager][nm-cfmu] and
computed consistently throughout the entire [Single European Sky Area][ses-area]
which allows comparisons across different units.

### Comparability over time
Comparisons over time are valid.

## Cost and burden

Not available. 

## Data revision

The data revision policy has not yet been defined. 

## Statistical processing

### Source data

Actual en route ATFM delay

* Data provided by the [Network Manager][nm-cfmu].

Actual Number of [IFR][ifr] flights

* Data provided by the [Network Manager][nm-cfmu].

### Data collection

Data is collected and compiled by the [Network Manager][nm-cfmu].

### Data validation

Data validation is the responsibility of the [Network Manager][nm-cfmu].

### Adjustment

Corrections may be made when issues are discovered as described in the
[Quality Assessment section of this document](#quality-assessment).

## Comment

No comments.

### Disclaimer {.disclaimer}

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'disclaimer.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


##

[no3128_2024]: https://eur-lex.europa.eu/legal-content/EN/TXT/?uri=CELEX:32024R3128 "IR 3128/2024"
[air-stats]: https://ansperformance.eu/definition/air-transport-statistics/ "Air Transport Statistics"
[ansp]: https://ansperformance.eu/acronym/ansp/ "ANSP"
[acc]: https://ansperformance.eu/acronym/acc/ "ACC"
[atfm-delay]: https://ansperformance.eu/definition/atfm-delay/ "ATFM Delay"
[coda]: https://www.eurocontrol.int/articles/central-office-delay-analysis-coda "CODA"
[ctot]: https://ansperformance.eu/acronym/ctot/ "CTOT"
[etot]: https://ansperformance.eu/acronym/etot/ "ETOT"
[fab]: https://ansperformance.eu/acronym/fab/ "FAB"
[fir]: https://ansperformance.eu/acronym/fir/ "FIR"
[fmp]: https://ansperformance.eu/acronym/fmp/ "FMP"
[kpi]: https://ansperformance.eu/acronym/kpi/ "KPI"
[ifr]: https://ansperformance.eu/acronym/ifr/ "IFR"
[rp1]: ../rp1/ "RP1"
[rp2]: ../rp2/ "RP2"
[rp3]: ../rp3/ "RP3"
[law]: https://ansperformance.eu/bibliography/legislation/ "Legislation"
[nm-cfmu]: https://ansperformance.eu/acronym/nm-cfmu/ "NM"
[nm20]: https://ansperformance.eu/library/NM_20_FB750_Delay_Calculation_improvements.pdf "REA Message"
[pru]: https://ansperformance.eu/about/us/ "PRU"
[refloc]: https://ansperformance.eu/definition/reference-location/ "Reference Location"
[regu]: https://ansperformance.eu/definition/regulation/ "Regulation"
[ses-area]: https://ansperformance.eu/definition/ses-area/ "SES Area"
[ses-portal]: ../../ "SES Data Portal"
[postops]: https://www.eurocontrol.int/service/post-operations-performance-adjustment "Post-operations performance adjustment"
