---
title: Flexible Use of Airspace (FUA) Performance Indicators 
categories:
  - metadata
  - glossary
---

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('static', 'css', 'table.css'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Contacts

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'contacts.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


## Metadata update

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'metadata-update.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```

### General

Set 1 – Effectiveness of ASM Process in terms of airspace made available 
Set 2 – Indicators on the utilisation of the available airspace 

Indicators are calculated in accordance with the provisions of IR(EU) 2019/317 ANNEX I, Section 1 points 2.2 (c),(d),(e), Section 2 points 2.2 (f),(g),(h) and Section 3 points 4.2 (e),(f),(g)

### Definition and details on calculation methodology
<a href="https://www.eurocontrol.int/sites/default/files/2021-07/ernip-part-3-asm-handbook-edition-5-ver-5-8.pdf" target="_blank">European Route Network Improvement Plan - Part 3 – ASM Procedures – Section 7</a>
