---
title: Minutes of en route ATFM delay per flight
categories:
  - metadata
  - glossary
---

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('static', 'css', 'table.css'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Contacts

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'contacts.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


## Metadata update

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'metadata-update.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


## Statistical presentation

### Data description

The measure provides an indication of [ATFM delays][atfm-delay] on the ground due to
constraints en-route.

In accordance with EU legislation, [Single European Sky Area][ses-area] States are required to
adopt performance targets contributing adequately to and be consistent with the EU wide targets.
The capacity [KPI][kpi] is the [*en route ATFM delay per flight*][atfm-delay].

In Europe, when traffic demand is anticipated to exceed the available capacity in en route centres
or at airports, Air Traffic Control (ATC) units may request the local
Flow Management Position ([FMP][fmp]) to instigate an Air Traffic Flow Management (ATFM) measure,
or regulation.
Aircraft expected to arrive during a period of congestion are given Air Traffic Flow Management
delay at their departure airport, under the authority of the [Network Manager][nm-cfmu],
in order to regulate the flow of traffic into the constrained downstream en route sector or airport,
thus ensuring safety.
The resulting ATFM delays are calculated as the difference between the estimated take-off time
calculated from the filed flight plan including updates and the calculated take-off time
allocated by the central unit of ATFM.
The reason for the regulation is indicated by the responsible Flow Management Position (FMP).
The delay is attributed to the FIR containing the reference location of the most constraining
capacity constraint.  

**En route ATFM delay** has been in use as a commonly agreed proxy for en route capacity shortfalls
since 1999.
The calculation of ATFM delays is based on a well established and commonly accepted algorithm
and is compliant with the [KPI][kpi] definition in the EU legislation.


### Classification system
En route ATFM delay per flight is classified per:

* [Single European Sky Area][ses-area] 
* [Functional Airspace Block][fab] 
* [Flight Information Region - FIR][fir]
* [Air Navigation Services Provider][ansp] 

It should be noted that the areas of responsibility of the individual [ANSP][ansp]s
are not necessarily coincidental with the geographical boundaries of the State.  

For the monitoring of National/FAB Performance targets, the [FAB][fab] is considered as
the aggregation of the individual [FIRs][fir] belonging to the Member state within the FAB.

### Sector coverage

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'sector-coverage.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


### Statistical concepts and definitions

ATFM delay
: *ATFM delay* is defined as the duration between the last Estimated Take-Off Time ([ETOT][etot])
  and the Calculated Take-Off Time ([CTOT][ctot]) allocated by the [Network Manager][nm-cfmu].
  ATFM delay comprises both Airport ATFM delay] and En route ATFM delay.
  This indicator is solely concerned with En route ATFM delay.
:  The delay is truncated to minutes.

Flights 
: All the [IFR][ifr] flights which have filed a flight plan and have entered the area of
  responsibility of the relevant [ANSP][ansp](s) are taken into account.
: Each flight is only counted once, regardless of its trajectory.

#### Attribution of Delay

[ATFM delay][atfm-delay] to which a flight is subjected is defined to have as cause
the most penalizing [ATFM regulation][regu] which affects it.
Depending of the type and the entity responsible for the [reference location][refloc] protected by
the [ATFM regulation][regu] the [ATFM delay][atfm-delay] is classified as  either an airport or
an en route [ATFM delay][atfm-delay] and attributed to the  [FIR][fir] containing the
[reference location][refloc].

Each ATFM regulation also comprises a designator referring to the cause of delay
which can be used in further analysis.

### Statistical unit

The statistical unit is the State/ANSP to which the ATFM delays are attributed.

### Statistical population

The statistical population comprises the States in the [Single European Sky Area][ses-area].

### Reference area

The reference area is the Single European Sky Area. 


### Time coverage

2012 is the first year for which data is presented.

### Base period

Not applicable.

## Unit of Measure

The en route [ATFM delay][atfm-delay] is measured in minutes per [IFR][ifr] flight.


## Reference period
```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'reference-period.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Institutional Mandate

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'institutional-mandate.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Confidentiality

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'confidentiality.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```

## Release policy

### Release calendar

En route [ATFM delay][atfm-delay] data is released 12 times per year, once per month.

### Release calendar access

Not applicable.

### User access

Information is disseminated to the general public via the [SES Data Portal][ses-portal]. 

## Frequency of dissemination

En route ATFM data is published monthly with the annual performance aggregate being available
in January of the following year.

## Dissemination format

Information is disseminated to the general public via the [SES Data Portal][ses-portal]. 

## Accessibility of documentation

### Documentation on methodology

As per the pertaining regulations (see [Institutional Mandate](#institutional-mandate)).<br>
For technical questions not addressed in [Statistical processing](#statistical-processing)
please contact: [PRU-Support@eurocontrol.int](mailto:PRU-Support@eurocontrol.int).<br>
Additional definitions of the terms used in the frame of the this [KPI][kpi]
are available in the [Metadata page](..).

### Quality documentation

There is no specific documentation on procedures applied for quality management
and quality assessment.

## Quality Management

### Quality assurance

The data is provided by the [Network Manager][nm-cfmu] and as such it is their
responsibility to perform any quality assurance activities.

### Quality assessment

Data is validated by random quality checks, e.g. missing data for given a day,
adherent data, etc.  Any issues are reported to the [Network manager][nm-cfmu] who works with the
[EUROCONTROL Performance Review Unit][pru] to resolve them.


## Relevance

The information is published for performance monitoring purposes in accordance with
the relevant EU legislation. 

### Completeness

The data is collected for the entire [Single European Sky Area][ses-area].

## Accuracy and reliability

### Overall accuracy

The en route [ATFM delay][atfm-delay] is the delay calculated by the [Network Manager][nm-cfmu].
This calculation is based on a well established methodology, which has been in place since 1999.

It is assumed that the data provided is accurate, given the caveats mentioned in sections
[Quality assurance](#quality-assurance) and
[Quality assessment](#quality-assessment).

### Sampling error

There is no sampling and therefore no sampling error. 

## Timeliness and punctuality

### Timeliness

The full set of Minutes of en route [ATFM delay][atfm-delay] per flight is published each month
- in general between 14 and 16 days after the end of the month in question.

###  Punctuality

The internal databases are updated daily. The [statistical processing](#statistical-processing)
is performed once per month.

## Comparability
### Comparability — geographical

The data is collected centrally by the [Network Manager][nm-cfmu] and
computed consistently throughout the entire [Single European Sky Area][ses-area]
which allows comparisons across different units.

### Comparability over time
Comparisons over time are valid.

## Coherence
### Coherence — cross domain

[ATFM delay][atfm-delay] is calculated relative to the flight plan. 

Although not fully cohered, the delay reported by aircraft operators to CODA
under the corresponding IATA delay categories shows a high correlation with the ATFM delay
calculated by the Network Manager. 
It should be noted that the delay reported by aircraft operators relates to airline schedules
rather than flight plan.

### Coherence — internal

Data is fully coherent from an internal perspective.

## Cost and burden

Not available. 

## Data revision

The data revision policy has not yet been defined. 

## Statistical processing

### Source data

Actual en route ATFM delay

* Data provided by the [Network Manager][nm-cfmu].

Actual Number of [IFR][ifr] flights

* Data provided by the [Network Manager][nm-cfmu].

Please note that, for operational purposes, the delays caused by READY to Depart (REA) messages 
and ATFM slot extensions were excluded from the ATFM delay computation with NM release 20 (4 April 2016) 
which reduces the overall amount of computed ATFM delay.
More information on the adjusted ATFM delay computation as of NM release 20.0
from 4 April can be found here: [NM release 20.0 - 4 April 2016 - FB750-Delay Calculation improvements][nm20].

### Frequency of data collection

Data is collected for each individual flight in real time.

### Data collection

Data is collected and compiled by the [Network Manager][nm-cfmu].

### Data validation

Data validation is the responsibility of the [Network Manager][nm-cfmu].

### Data compilation 

The data can be aggregated at different levels. 

| Reference period | EU wide | Local                   |
|------------------|---------|-------------------------|
| [RP1][rp1] | SES RP1 area | [Functional Airspace Block][fab], [Air Navigation Services Provider][ansp] |
| [RP2][rp2] | SES RP2 area | [Functional Airspace Block][fab], Country [FIR][fir]|
| [RP3][rp3] | SES RP3 area | [Functional Airspace Block][fab], [Air Navigation Services Provider][ansp] |

### Adjustment

Corrections may be made when issues are discovered as described in the
[Quality Assessment section of this document](#quality-assessment).

#### Post-operations performance adjustment

The post-operations performance adjustment process allows operational stakeholders to notify national and European 
authorities of problems related to air traffic flow management (ATFM) delay measurement, classification and assignment.
The main output of this process is a separate performance dataset, which will include the approved changes. 
This dataset is available for performance scheme and local management reporting.
[More information on the post-operations performance adjustment process can be found here][postops] 

## Comment

No comments.

### Disclaimer {.disclaimer}

```{r, echo=FALSE, results='asis'}
res <- knitr::knit_child(
  here::here('parts', 'disclaimer.Rmd'),
  quiet = TRUE)
cat(res, sep = '\n')
```


##






[air-stats]: https://ansperformance.eu/definition/air-transport-statistics/ "Air Transport Statistics"
[ansp]: https://ansperformance.eu/acronym/ansp/ "ANSP"
[atfm-delay]: https://ansperformance.eu/definition/atfm-delay/ "ATFM Delay"
[coda]: https://www.eurocontrol.int/articles/central-office-delay-analysis-coda "CODA"
[ctot]: https://ansperformance.eu/acronym/ctot/ "CTOT"
[etot]: https://ansperformance.eu/acronym/etot/ "ETOT"
[fab]: https://ansperformance.eu/acronym/fab/ "FAB"
[fir]: https://ansperformance.eu/acronym/fir/ "FIR"
[fmp]: https://ansperformance.eu/acronym/fmp/ "FMP"
[kpi]: https://ansperformance.eu/acronym/kpi/ "KPI"
[ifr]: https://ansperformance.eu/acronym/ifr/ "IFR"
[rp1]: ../rp1/ "RP1"
[rp2]: ../rp2/ "RP2"
[rp3]: ../rp3/ "RP3"
[law]: https://ansperformance.eu/bibliography/legislation/ "Legislation"
[nm-cfmu]: https://ansperformance.eu/acronym/nm-cfmu/ "NM"
[nm20]: https://ansperformance.eu/library/NM_20_FB750_Delay_Calculation_improvements.pdf "REA Message"
[pru]: https://ansperformance.eu/about/us/ "PRU"
[refloc]: https://ansperformance.eu/definition/reference-location/ "Reference Location"
[regu]: https://ansperformance.eu/definition/regulation/ "Regulation"
[ses-area]: https://ansperformance.eu/definition/ses-area/ "SES Area"
[ses-portal]: ../../ "SES Data Portal"
[postops]: https://www.eurocontrol.int/service/post-operations-performance-adjustment "Post-operations performance adjustment"
